\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-26 forwards-and-backwards.tex
symlinks to figures
alison edits

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        \includegraphics{figures/ex1b/fkpp_123.density.3.0}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
    \end{center}
    \caption{
        Simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_1d/clumpy_123.steps}
    \end{center}
    \caption{
        \textbf{Left:} A snapshot of individual locations in a two-dimensional simulation
        in which the constant density is unstable
        and a stable, periodic pattern forms.
        \textbf{Right:} Population density in an expanding wave in a one-dimensional simulation
        forming a periodic pattern;
        each panel shows the wavefront in three periods of time;
        within each period of time the wavefront at earlier times is shown in blue
        and later times in pink.
        In both cases, $\gamma(m) = 3/(1 + m)$, $\mu \equiv 0.3$, and $r \equiv 1$;
        dispersal is Gaussian with $\sigma=0.2$ and density is measured with $\rho_\gamma(x) = p_9(x)$,
        i.e., using a Gaussian kernel with standard deviation 3.
        \label{fig:clumping}
    }
\end{figure}

@@

\comment{TODO: check that $\eta_s^N$ having an $N$ or not is okay}

@@

The first part of the proof mirrors that of Theorem~\ref{thm:nonlocal_convergence}:
we establish sufficient bounds on the moments of $\rho_\epsilon*\eta_t(x)$ to
establish tightness and then apply standard results on 
convergence of Markov processes from~\cite{ethier/kurtz:1986}. 
The challenge
comes in identifying the limit points. This is much more intricate than the case in 
which we do not scale the interaction kernel, as weak convergence will no longer
be sufficient to guarantee the form of the nonlinear terms in the limiting equation.
Identification of the limit will rest on regularity inherited from continuity 
estimates for a random walk with Gaussian jumps which we prove in 
Subsection~\ref{continuity for random walk}.
Various bounds on the moments of $\rho_\epsilon*\eta_t$, one consequence of 
which is tightness, are then provided in 
Subsection~\ref{bounds on rhoepsilon},
and we identify the limit points in Subsection~\ref{limit in onestep case}.

\subsection{Continuity estimates for a Gaussian random walk}
\label{continuity for random walk}

@@

For the second claim, first define $\tau(t) = T(t)-t$.
Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
then partitioning over
$\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
\begin{align}
   \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{\big(2 \pi (\epsilon^2+T(t))\big)^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
\end{align}
Now, observe that since $\IE[e^{-\Pi(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
by Markov's inequality,
\begin{align}
    \IP\left\{
        \tau(t) < - \frac{\epsilon^2 + t}{2}
    \right\}
    &=
    \IP\left\{
        e^{-\Pi(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
    \right\}
\nonumber
\\
&\leq \frac{\IE[\exp\big(-\Pi(\theta t)\big)]}{\exp\big(-\theta(t-\epsilon^2)/2\big)}
\nonumber
    \\&=
    \frac{
        \exp(- \theta t(1 - e^{-1}))
    }{
        \exp(- \theta (t - \epsilon^2) / 2)
    }
\nonumber
    \\&=
    \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
\label{bound for negative tau}
\end{align}
where $\delta = 1/2 - e^{-1} > 0$.
The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
\[
    C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t - \epsilon^2 \theta / 2}
    \frac{1}{(\epsilon^2 + t)^{d/2}} .
\]
Now observe that the derivative (with respect to $t$) of 
$e^{-\delta \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ 
is
\[
    \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
    \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t - \epsilon^2 \theta / 2} ,
\]
which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
At the maximum, $(\epsilon^2 + t) = d \delta / (2 \theta)$,
so $(1 + t/\epsilon^2) = d \delta/(2 \theta \epsilon^2)$,
and so this quantity is bounded uniformly over not only $t$ but also $\epsilon$
(since we've assumed that $\theta \epsilon^2$ is bounded below).
Therefore, the second term in \eqref{eqn:psi_infty_bound} 
is bounded by $C / (\epsilon^2 + t)^{d/2}$
(for a different constant $C$).
Substituting this into \eqref{eqn:psi_infty_bound} yields the result.
\end{proof}

We are going to approximate $\psi_t^{\epsilon,x}(\cdot)$ by $p_{\epsilon^2+t}(x, \cdot)$,
and to control the error that this introduces we 
need to control $T(t)-t$. We achieve this through a large deviations bound.

\begin{lemma}
    \label{lem:poisson_ld}
    Fix $A > e^3 - 1$ (so that $\log(1 + A) > 3$).
    In the notation of Lemma~\ref{PsiBoundHS},
    \begin{align*}
        \limsup_{\theta \to \infty}
        \frac{1}{\theta} \log
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
    \le
        - (\epsilon^2 + t) \frac{A}{2}
        - \log(1 + A)\left( \frac{t + A(\epsilon^2 + t)}{2} \right) .
    \end{align*}
\end{lemma}

\begin{proof}
    We apply Cram\'er's Theorem.
    First note that
    \begin{align*}
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
        =
        \IP\Big\{
            \frac{\Pi(\theta t)}{ \theta}
            >
            t + A(\epsilon^2 + t)
        \Big\} .
    \end{align*}
For convenience, suppose $\theta$ is an integer,
and so we may write $\Pi(\theta t) = Z_1(t) + \cdots + Z_\theta(t)$,
where the $Z_i(t)$ are i.i.d.~Poisson($t$) random variables.
    To apply Cram\'er's Theorem, we need the rate function for the 
Poisson($t$) distribution,
    which is
    \[
        \Lambda^*(x) = \sup_{\alpha > 0} \left[ \alpha x - \lambda(\alpha) \right],
    \]
    where $\lambda(\alpha) = \log \IE[\exp(\alpha Z_1(t))] = t(e^\alpha - 1)$.
    Now, the supremum of $\alpha x + t - te^\alpha$ is attained at $x = t e^\alpha$,
    and so
    \[
        \Lambda^*(x) = x \log(x/t) - x + t .
    \]
    Cram\'er's Theorem then gives
    \begin{align*}
        \limsup_{\theta \to \infty}
        \frac{1}{\theta} \log
        \IP\left\{
            \frac{\Pi(\theta t)}{\theta}
            > t + A(\epsilon^2 + t)
        \right\}
    &\le
        - \Lambda^*(t + A(\epsilon^2 + t))
    \\ &=
        A(\epsilon^2 + t)
        -
        \big(t + A(\epsilon^2 + t)\big) \log\left( 1 + \frac{A(\epsilon^2 + t)}{t} \right) 
    \\ &\le
        A(\epsilon^2 + t)
        -
        \big(t + A(\epsilon^2 + t)\big) \log(1 + A)
    \\ &\le
        A(\epsilon^2 + t) \left( 1 - \frac{1}{2} \log(1 + A) \right)
        -
        \frac{1}{2} \big(t + A(\epsilon^2 + t)\big) \log(1 + A)  .
    \end{align*}
    Since $(1/2) \log(1 + A) > 3/2$, from which $1 - (1/2) \log(1 + A) < - 1/2$,
    the proof is complete.
\end{proof}

@@

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    and with $A$ as in Lemma~\ref{lem:poisson_ld},
    \begin{multline}
\label{heat kernel estimate}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \\
\le
        \frac{C}{(\epsilon^2 \theta)^{d/2}}
        p_{2(A+1)(\epsilon^2+t)}(x, y)
        +
        C \exp(- \epsilon^2 \theta / 2)
        +
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \IP\left\{
            \tau(t) > A(\epsilon^2 + t)
        \right\} .
    \end{multline}
\end{lemma}

>>>>

\begin{figure}
    \begin{center}
        % \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        % \includegraphics{figures/ex1b/fkpp_123.density.3.0}
        \includegraphics{figures/fig1a}
        \includegraphics{figures/fig1b}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=0.7\textwidth]{figures/fig2a}
        \includegraphics[width=0.7\textwidth]{figures/fig2b}
    \end{center}
    \caption{
        Simulated 1D populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        % \includegraphics[width=0.45\textwidth]{figures/clumpy_1d/clumpy_123.steps}
        \includegraphics[width=0.45\textwidth]{figures/fig3a}
        \includegraphics[width=0.45\textwidth]{figures/fig3b}
    \end{center}
    \caption{
        \textbf{Left:} A snapshot of individual locations in a two-dimensional simulation
        in which the constant density is unstable
        and a stable, periodic pattern forms.
        \textbf{Right:} Population density in an expanding wave in a one-dimensional simulation
        forming a periodic pattern;
        each panel shows the wavefront in three periods of time;
        within each period of time the wavefront at earlier times is shown in blue
        and later times in pink.
        In both cases, $\gamma(m) = 3/(1 + m)$, $\mu \equiv 0.3$, and $r \equiv 1$;
        dispersal is Gaussian with $\sigma=0.2$ and density is measured with $\rho_\gamma(x) = p_9(x)$,
        i.e., using a Gaussian kernel with standard deviation 3.
        \label{fig:clumping}
    }
\end{figure}

@@

@@

The first part of the proof mirrors that of Theorem~\ref{thm:nonlocal_convergence}:
we establish sufficient bounds on the moments of $\rho_\epsilon*\eta_t(x)$ to
establish tightness and then apply standard results on 
convergence of Markov processes from~\cite{ethier/kurtz:1986}. 
Bounds on the moments of $\rho_\epsilon*\eta_t$
are established in 
Subsection~\ref{bounds on rhoepsilon}.
The challenge
comes in identifying the limit points. This is much more intricate than the case in 
which we do not scale the interaction kernel, as weak convergence will no longer
be sufficient to guarantee the form of the nonlinear terms in the limiting equation.
Identification of the limit will rest on regularity inherited from continuity 
estimates for a random walk with Gaussian jumps which we prove in 
Subsection~\ref{continuity for random walk}, before
identifying the limit points in Subsection~\ref{limit in onestep case}.

@@

For the second claim, first define $\tau(t) = T(t)-t$.
Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
partitioning over
$\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
\begin{align}
   \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{\big(2 \pi (\epsilon^2+T(t))\big)^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\}. \label{eqn:psi_infty_bound}
\end{align}
Now, observe that since $\IE[e^{-\Pi(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
by Markov's inequality,
\begin{align}
    \IP\left\{
        \tau(t) < - \frac{\epsilon^2 + t}{2}
    \right\}
    &=
    \IP\left\{
        e^{-\Pi(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
    \right\}
\nonumber
\\
&\leq \frac{\IE[\exp\big(-\Pi(\theta t)\big)]}{\exp\big(-\theta(t-\epsilon^2)/2\big)}
\nonumber
    \\&=
    \frac{
        \exp(- \theta t(1 - e^{-1}))
    }{
        \exp(- \theta (t - \epsilon^2) / 2)
    }
\nonumber
    \\&=
    \exp\left\{ - \chi \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
\label{bound for negative tau}
\end{align}
where $\chi = 1/2 - e^{-1} > 0$.
The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
\[
    C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\chi \theta t - \epsilon^2 \theta / 2}
    \frac{1}{(\epsilon^2 + t)^{d/2}} .
\]
Now observe that the derivative (with respect to $t$) of 
$e^{-\chi \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ 
is
\[
    \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \chi \theta \right)
    \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\chi \theta t - \epsilon^2 \theta / 2} ,
\]
which is negative if $\theta(\epsilon^2 + t) > d/2 \chi$.
At the maximum, $(\epsilon^2 + t) = d / (2\chi \theta)$,
so $(1 + t/\epsilon^2) = d /(2\chi \theta \epsilon^2)$,
and so this quantity is bounded uniformly over not only $t$ but also $\epsilon$
(since we've assumed that $\theta \epsilon^2$ is bounded below).
Therefore, the second term in \eqref{eqn:psi_infty_bound} 
is bounded by $C / (\epsilon^2 + t)^{d/2}$
(for a different constant $C$).
Substituting this into \eqref{eqn:psi_infty_bound} yields the result.
\end{proof}

Now suppose that we have established~(\ref{moment bound rhoepsilon}) 
and~(\ref{flandoli trick for rhoepsilon}) for all exponents $j<k$.
First we calculate the generator $\Pgen^N$ of our scaled population process applied
to functions of the form $\langle f,\eta\rangle^k$.
Recalling that each jump of the process involves the birth or death or a single 
individual, and so increments $\langle f,\eta\rangle$ by $\pm f/N$ at the location of
that individual, we find 
\begin{multline}
\label{pgen applied to kth moment}
\Pgen^N\Big(\langle f,\eta\rangle^k\Big)
=
\Big\langle
\int\theta N\Big\{k\frac{f(y)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(y)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{f(y)^{k}}{N^{k}}\Big\}q_\theta(x,dy),\eta(dx)\Big\rangle
\\
+
\Big\langle\theta N\Big(1-\frac{F(\rho_\epsilon*\eta(x))}{\theta}\Big)
\Big\{-k\frac{f(x)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(x)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{(-1)^{k}f(x)^{k}}{N^{k}}\Big\},\eta(dx)\Big\rangle.
\end{multline}
Mimicking what we did above, we set $f(\cdot)=\psi^{\epsilon,x}_t(\cdot)$ and 
write
\begin{multline}
\label{kth moment for time varying function}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
=\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\IE\Big[\int_0^t\Pgen^N\big(\langle\psi^{\epsilon,x}_{t-s}(y),\eta_s(dy)\rangle^k\big) ds
\\
-\int_0^t\langle k\dot{\psi}^{\epsilon,x}_{t-s}(y), \eta_s(dy)\rangle
\big\langle \psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\big\rangle^{k-1}ds
\Big|{\cal F}_0\Big].
\end{multline}
Using(from Lemma~\ref{PsiBoundHS}) 
that $\sup_s\|\psi^{\epsilon,x}_s(\cdot)\|_\infty=C/\epsilon^d$,
$N\epsilon^d>1$,
and our inductive
hypothesis, we find
\begin{multline*}
\IE\Big[
\Big\langle
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\int
\frac{\psi_{t-s}^{\epsilon,x}(z)^j}{N^j}
\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
q_\theta(y,dz),
\eta_s(dy)\Big\rangle ds 
\\
+
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\frac{\psi_{t-s}^{\epsilon,x}(y)^j}{N^j}\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
(-1)^j\Big(1-\frac{F(\rho_\epsilon*\eta_s(y))}{\theta} 
\Big),
\eta_s(dy)\Big\rangle ds\Big|{\cal F}_0\Big]
\\
\leq 
C\IE\Big[\Big\langle\int_0^t\sum_{j=2}^k\frac{\theta}{N\epsilon^d}
\Big(\frac{1}{(N\epsilon^d)^{j-2}}\Big)
	\Big\langle \psi^{\epsilon,x}_{t-s}(y)
\Big(2+\frac{|F(\rho_\epsilon*\eta_s(y))|}{\theta}\Big),\eta_s(dy)\Big\rangle
\langle\psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}ds\big|{\cal F}_0\Big]
\\
\leq
C'\frac{\theta}{N\epsilon^d}\sum_{j=1}^{k-1}
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^j
\leq
C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big).
\end{multline*}
Combining this with~(\ref{pgen applied to kth moment})
and~(\ref{kth moment for time varying function}), using once again the fact that $F$
is bounded above, we find 
\begin{multline*}
\label{kth step of induction}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
\leq
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\tilde{C}\IE\Big[\int_0^t
\langle\psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\rangle^k ds
\Big|{\cal F}_0\Big]
\\
+C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k+
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big)
,
\end{multline*}
and~(\ref{moment bound rhoepsilon}) 
follows from Gronwall's inequality. Rearranging exactly as in the case $k=1$, we 
recover~(\ref{flandoli trick for rhoepsilon}) 
and the inductive step is complete.
\end{proof}

\begin{corollary}[Tightness of $\{(\rho_\epsilon*\eta_t^N(x) dx)_{t\geq 0}\}$]
\label{lem:density_tightness} 
Under the assumptions of Theorem~\ref{thm:local_convergence},
the sequence of measure valued processes $\{ \rho*\eta_t^N(x) dx \}_{t \geq 0}$
	(taking values in ${\mathcal D}_{[0,T]}(\measures)$) is tight.
\end{corollary}
\begin{proof}

By the Intermediate Value Theorem, using $T_t$ to denote the heat semigroup,
there exists $s\in (0,1/\theta)$ such that
\begin{multline*}
\left|
\theta\int\big(\rho_\epsilon*f(y)-\rho_\epsilon*f(x)\big)
q_{\theta}(x,dy) \right|
=
\left|\theta\Big(T_{\epsilon^2+1/\theta}f(x)-T_{\epsilon^2}f(x)\Big)\right|
\\
=
\left|\partial_sT_{\epsilon^2+s}f(x)\right|=\left|T_{\epsilon^2+s}\Delta f(x)\right|
\leq\|\Delta f\|_\infty.
\end{multline*}
We therefore have
\begin{multline*}
        \left|
        \langle \rho_{\epsilon} * f, \eta_t^N \rangle
        -
        \langle \rho_{\epsilon} * f, \eta_s^N \rangle
        \right|
        \leq
        \| \Delta f \|_\infty |t-s|
        \sup_{0 \leq u \leq T } \langle 1, \eta_u^N \rangle
\\
+
	\int_s^t
	\big\langle \rho_\epsilon*|f|(x)\, \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du +2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|.
    \end{multline*}
Using the Burkholder-Davis-Gundy inequality, our
moment bounds, and Markov's inequality
\begin{equation}
\label{martingale term to zero}
\limsup_{N\to\infty}\IP\Big[
2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|
>\frac{\nu}{3}
\Big]\leq \limsup_{N\to\infty}\frac{36}{\nu^2}\IE\big[[\widehat{M}^N(f)]_T\big]
\leq \limsup_{N\to\infty}\frac{36}{\nu^2}\frac{C\theta}{N\epsilon^d} 
=0. 
\end{equation}
Now consider
\begin{multline}
	\label{square of F-integral}
\IE\Big[
	\Big(\int_s^t\big\langle\rho_\epsilon*|f|(x) \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du\Big)^2\Big]
\\
=
2\IE\Big[
	\int_s^t\big\langle\rho_\epsilon*|f|(x) \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle 
	\int_u^t\big\langle \rho_\epsilon*|f|(x)\big|F\big(\rho_\epsilon*\eta_r^N(x)\big)\big|,
\eta_r^N(dx)\big\rangle dr
du\Big].
\end{multline}
Since $F$ is polynomial, we use the approach
of Corollary~\ref{alternative form moment bounds}, the tower 
property, and 
Lemma~\ref{bounds on moments}, to bound this
in terms of sums of terms of the form
\[
\IE\Big[\int_s^t(t-u)\int\rho_\epsilon*|f|(x)\rho_\epsilon*\eta_u^N(x)^jdx
\int\rho_\epsilon*|f|(y)\rho_\epsilon*\eta_u^N(y)^kdydu\Big].
\]
Now observe that, again using Lemma~\ref{bounds on moments}, 
\begin{multline*}
	\IE\left[	\int\int\rho_\epsilon*|f|(x)\rho_\epsilon*\eta_u^N(x)^j
	\rho_\epsilon*|f|(y)\rho_\epsilon*\eta_u^N(y)^k dx dy\right]
	\\
	\leq
	\IE\left[\int\int\|f\|_\infty\rho_\epsilon*\eta_u^N(x)^{j+k}
	\rho_\epsilon*|f|(y)
	dx dy
+
	\int\int
	\rho_\epsilon*|f|(x)
	\|f\|_\infty
	\rho_\epsilon*\eta_u^N(y)^{j+k}
	dx dy \right]
	\\ \leq C\int |f|(x)dx.
\end{multline*}
Thus the quantity~(\ref{square of F-integral}) 
is bounded by $C(t-s)^2$ for some constant $C$ which we can 
take to be independent of $s$, $t$ and $\epsilon$. Markov's inequality 
then gives
\[
\IP\Big[\|f\|_\infty\int_s^t
\big\langle \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du \geq \frac{\nu}{3}\Big]
\leq C\frac{(t-s)^2}{\nu^2}.
\]
A union bound gives that
\begin{equation}
\label{union bound}
\IP\Big[\max_i\|f\|_\infty\int_{t_{i-1}}^{t_i}
\big\langle \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du \geq \frac{\nu}{3}\Big]
\leq C\frac{T\delta}{\nu^2}.
\end{equation}
Now using Markov's inequality, we can choose $K$ so that
\[
\IP\Big[\|\Delta f\|_\infty\, \sup_{0\leq t\leq T} \langle 1,\eta_t^N\rangle 
>K\Big]<\frac{\nu}{3}, 
\]
and so choosing $\delta$ so that $K\delta<\nu/3$ in this 
expression and $C\delta<\nu^3/3T$ in~(\ref{union bound}),
combining with~(\ref{martingale term to zero}), the result follows.
\end{proof}

\subsection{Continuity estimates for $\rho_\epsilon*\eta$}
\label{continuity for random walk}

In order to identify the limit point of any convergent subsequence
of $\{\rho_\epsilon*\eta^N(x)\}$, 
we are going to require some control on the spatial continuity of the 
functions $\rho_\epsilon*\eta^N(x)$. This will be inherited from the regularity
of the transition density of the 
Gaussian random walk with generator ${\cal L}^\theta$, which in turn follows
from its representation
as that of a Brownian motion evaluated at the random time $T(t)$ defined
in Lemma~\ref{PsiBoundHS}.
Our approach will be to 
approximate $\psi_t^{\epsilon,x}(\cdot)$ by $p_{\epsilon^2+t}(x, \cdot)$,
and to control the error that this introduces we 
need to control $T(t)-t$.

\begin{lemma}
    \label{lem:poisson_ld}
    In the notation of Lemma~\ref{PsiBoundHS}, for any $A>1$,
    \begin{align*}
        %\limsup_{\theta \to \infty}
        %\frac{1}{\theta} \log
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
    \le
	    \exp\left(-\frac{\theta A}{4}\big(\epsilon^2+t\big)
	    %- (\epsilon^2 + t) \frac{A}{2}
        %- \log(1 + A)\left( \frac{t + A(\epsilon^2 + t)}{2} 
	    \right) .
    \end{align*}
\end{lemma}

\begin{proof}
    %We apply Cram\'er's Theorem.
    %First note that
    This is just a Chernoff bound. With $\Pi$ a rate one Poisson process as
	in Lemma~\ref{PsiBoundHS}, for any $A>1$,
	\begin{align*}
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
&=
        \IP\Big\{
            \Pi(\theta t)
            >
		\theta\Big( t + A(\epsilon^2 + t)\Big)
        \Big\} \\
&\leq\frac{\IE\left[\exp\left(\alpha\Pi(\theta t)\right)\right]}
		{\exp\left(\alpha\theta\big(t+A(\epsilon^2+t)\big)\right)}
		\\
		&=\exp\left(\theta t\big(e^\alpha-1\big)-\alpha\theta\big(t+A(\epsilon^2+t)\big)\right)
		\\
		&\leq
		\exp\left(\theta t\big(e^\alpha-\alpha-1-\frac{A\alpha}{2}\big)
		-\frac{A\alpha}{2}\theta(\epsilon^2+t)\right).
    \end{align*}
	Now set $\alpha=1/2$. Since $A>1$, $e^\alpha-\alpha-1-A\alpha/2<0$ and the 
	result follows.
%For convenience, suppose $\theta$ is an integer,
%and so we may write $\Pi(\theta t) = Z_1(t) + \cdots + Z_\theta(t)$,
%where the $Z_i(t)$ are i.i.d.~Poisson($t$) random variables.
%    To apply Cram\'er's Theorem, we need the rate function for the 
%Poisson($t$) distribution,
%    which is
%    \[
%        \Lambda^*(x) = \sup_{\alpha > 0} \left[ \alpha x - \lambda(\alpha) \right],
%    \]
%    where $\lambda(\alpha) = \log \IE[\exp(\alpha Z_1(t))] = t(e^\alpha - 1)$.
%    Now, the supremum of $\alpha x + t - te^\alpha$ is attained at $x = t e^\alpha$,
%    and so
%    \[
%        \Lambda^*(x) = x \log(x/t) - x + t .
%    \]
%    Cram\'er's Theorem then gives
%    \begin{align*}
%        \limsup_{\theta \to \infty}
%        \frac{1}{\theta} \log
%        \IP\left\{
%            \frac{\Pi(\theta t)}{\theta}
%            > t + A(\epsilon^2 + t)
%        \right\}
%    &\le
%        - \Lambda^*(t + A(\epsilon^2 + t))
%    \\ &=
%        A(\epsilon^2 + t)
%        -
%        \big(t + A(\epsilon^2 + t)\big) \log\left( 1 + \frac{A(\epsilon^2 + t)}{t} \right) 
%    \\ &\le
%        A(\epsilon^2 + t)
%        -
%        \big(t + A(\epsilon^2 + t)\big) \log(1 + A)
%    \\ &\le
%        A(\epsilon^2 + t) \left( 1 - \frac{1}{2} \log(1 + A) \right)
%        -
%        \frac{1}{2} \big(t + A(\epsilon^2 + t)\big) \log(1 + A)  .
%    \end{align*}
%    Since $(1/2) \log(1 + A) > 3/2$, from which $1 - (1/2) \log(1 + A) < - 1/2$,
%    the proof is complete.
\end{proof}

@@

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    %and with $A$ as in Lemma~\ref{lem:poisson_ld},
    \begin{multline}
\label{heat kernel estimate}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \\
\le
        \frac{C}{(\epsilon^2 \theta)^{d/2}}
        p_{6(\epsilon^2+t)}(x, y)
        +
	    C \left(1+
	    \frac{1}{(\epsilon^2 + t)^{d/2}}\right)
	    \exp(- \epsilon^2 \theta / 2)
        %+
        %\frac{C}{(\epsilon^2 + t)^{d/2}}
        %\IP\left\{
        %    \tau(t) > A(\epsilon^2 + t)
        %\right\} 
	    .
    \end{multline}
\end{lemma}
